
\documentclass{article}

\usepackage[hmargin=3cm,vmargin=3.5cm]{geometry}
\usepackage{enumerate,bm}

\begin{document}

\begin{center}
{\large Homework 1: BIOS 623}\\
{\large Due September 12 by 11:59pm}
\end{center}

Complete all questions.  All homework solutions should be typed, or at least in soft copy. \\


\begin{enumerate}[Q 1.]
\item HW\_Q1.sas7bdat gives the systolic blood pressure (SBP), body size (QUET), age (AGE), and smoking history ($SMK =0$ if nonsmoker, $SMK=1$ if current or previous smoker) for a hypothetical sample of 32 white males over 40 years old from the town of Angina.
	\begin{enumerate}[a)]
	\item Completely state the model needed to regress QUET (X) on SBP (Y).
	\item Determine the least-squares estimates of the slope ($\beta_1$) and the intercept ($\beta_0$) for the straight-line regression of SBP (Y) on QUET (X).
	\item Perform the test of $H_0:~\beta_1 = 0$, including all relevant items for hypothesis testing.
	\item Plot the line-of-best-fit along with 95\% prediction bands.
	%\item Fine an appropriate 95\% prediction interval for an individual with QUET=3.4 (the sample mean of QUET).  Interpret your answer.
	\end{enumerate}
\item Again using HW\_Q1.sas7bdat:
	\begin{enumerate}[a)]
	\item Completely state the model needed to regress SMK (X) on SBP (Y).
	\item Determine the least-squares estimates of the slope and intercept for the straight-line regression of SBP (Y) on SMK (X).
	\item Compare the value of $\hat\beta_0$ with the sample mean SBP for nonsmokers.  Compare the values of $\hat\beta_0 + \hat\beta_1$ with the mean SBP for smokers.  Describe the results of these comparisons.
	\item Perform the test of $H_0:~\beta_1 = 0$, including all relevant items for hypothesis testing.
	\item Perform the two-sample $t$ test for the equality of two population means with equal variance to compare the average SBP between Smokers and Non-smokers. 
	\item Are the tests in parts d) and )e equivalent? Provide theoretical justification of your answer, that is don't just say `they are the same' or `they are different.' Prove it. 
	\end{enumerate}
\item HW\_Q2.sas7bdat presents the starting annual salaries (SAL) of a group of 30 college graduates who have recently entered the job market, along with their cumulative grade-point averages (GPA).
	\begin{enumerate}[a)]
	\item Completely state the model needed to regress GPA (X) on SAL (Y).
	\item Determine the least-squares estimates of the slope and intercept of for the straight-line regression of SAL (Y) on GPA (X). Include a plot of the line-of-best fit with 95\% Confidence and Prediction bands.
	%\item Are any of the assumptions for straight-line regression clearly not satisfied in this example?
	\item Obtain a 95\% confidence interval for $\beta_1$. Would you reject the null hypothesis $H_0:~\beta_1 = 4000$ at the $\alpha=0.05$ level?
	\item Would you reject the null hypothesis $H_0:~ \mu_{Y|X} = 11,500$ at $X=2.75$?
	\end{enumerate}
	
	\newpage
	\clearpage

\item Let $\bm{A}$ = 
$\left[ \begin{array}{ccc} 
1 & 2 & 3\\ 
2 & 1 & 4 \\ 
3 & 4 & 1 
\end{array} \right] 
,~ \bm{B} = 
\left[ \begin{array}{cc} 
1 & 2 \\ 
2 & 1  
\end{array} \right] 
\mbox{, and }  \bm{C} = 
\left[ \begin{array}{cc} 
-1/3 & 2/3 \\ 
2/3 & -1/3  
\end{array} \right] . $
\begin{enumerate}[a)]
	\item What are the dimensions of $\bm{A}$, $\bm{B}$, and $\bm{C}$? 
	\item What is the result of $\bm{A} + \bm{B}$?
	\item What is the result of $\bm{B}*\bm{C}$?  What kind of matrix is $\bm{B}*\bm{C}$?
	\end{enumerate}

\item Prove by hand the following showing all necessary steps:
	\begin{itemize}
		\item $SST = SSM + SSE$
		\item $\sum_{i=1}^n e_i = 0$
		\item $Var({\bm{\hat\beta}}) = \sigma^2(\bm{X}'\bm{X})^{-1}$
	\end{itemize}
	
%\item What is the definition of a p-value? In your opinion, what are the pros/cons of the common significance level of 0.05 for p-values?
	
\item Read ``The Nutritional and Social Environment-Related Effects of Breastfeeding on Intelligence'' by Kosse. Answer the following questions:
	\begin{itemize}
		\item What is the population used to perform the analysis"?
		\item Kosse postulates that the average IQ at 0 weeks is ``significantly higher'' than the predicted value from the proposed models.  Can we confirm that from the data provided? If not, what statistical information would you like to see to confirm that statement?
		\item Do you believe that this paper should have been published?
	
	\end{itemize}
	
\end{enumerate}

\end{document}